Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
Overview of sec 5

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{??}).
However, not all $\Ltms$ programs are \gls{sms}.\MP{reworded. was imprecise}
That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{??}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{??}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
\MPin{
	no, it's different. 
	there is no CCT because we asume it's out of the programmer's concers and since CT is arch-dependent, the compiler should preserve CT, not the programmer, via software
}

Language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{??}).
However, $\Lscct$ also contains a countermeasure to enforce \gls{cct}, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
\MPin{ref to what real world stuff justifies this}

>>>>

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

This paper considers \gls{cct} as an architecture-dependent property that should not be one of the programmer's concerns~\cite{cauligi2019fact}.
Such consideration is backed up by architecture providing a data (operand) independent timing mode, such as processors by Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman}.
In spirit of this, the language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{subsec:lscct}), but provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
